Sentiment classification of online political discussions: a comparison of a word-based and dependency-based method

### 2014-03-21 manuscript.tex
Changed manuscript

\title{Sentiment classification of online political discussion -- a comparison of two methods}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

@@

%% \end{center}
\maketitle
\begin{abstract}
Bla bla
\end{abstract}

@@

Over the past years there has been an alarming growth in hate against minorities like Muslims, Jews, Gypsies and gays in Europe, driven by right wing populism parties and extremist organizations \cite{r4,r11}. A similar increase in hate speech has been observed on the Internet \cite{r6,s2}, and experts are concerned that individuals influenced by this web content may resort to violence as a result \cite{Strommen12,Sunde13}. Hateful speech is not only observed on extremist sites, but also as comments on e.g. Twitter, YouTube and online newspaper articles.

Social media and online discussions contain a wealth of information which can make us able to understand the extent of hate speech on the Internet. However, it turns out that academia is lacking research on social media and online radicalization \cite{s1}. Sentiment analysis (SA) is the discipline of automatically determining sentiment in text material and may be one important tool in the understanding of online radicalization. SA has mostly been used to analyze opinions in comments and reviews about commercial products, but there are also examples of SA towards political tweets and discussions, see e.g. Tumasjan et al. \shortcite{Tumasjan2010}; Chen et al. \shortcite{Chen10}. SA of political discussions is known to be a difficult task since citations, irony and sarcasm is very common \cite{Bing12}.

Sentiment classification aims to classify a document or sentence as either positive or negative and sometimes also neutral. There are mainly two approaches, one based on machine learning and one based on based on using a list of words with given sentiment scores (lexicon-based approach). One simple lexicon-based approach is to count the number of words with positive and negative sentiment in the document as suggested by Hu and Liu \shortcite{Hu04}. One may classify the opinion of larger documents like movie or product reviews or smaller documents like tweets, comments or sentences. %% See Liu \shortcite{Bing12}, chapters three to five and references therein for the description of several opinion classification methods.

@@

The paper is organized as follows.

@@

In this section we present two methods to classify sentences as either positive, neutral or negative towards a target word. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both target words, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the target word in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(kw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the target word appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters are within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the sentiment towards the target word to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set, as described in section \ref{sec:exp} below.

@@

For the word distance method we use the following \textbf{imp} function
\begin{equation}
  \label{eq:2}
  \mathbf{imp}(kw_i, sw_{j}) = \frac{\mathbf{sentsc}(sw_{j})}{\mathbf{worddist}(kw_i, sw_{j})}
\end{equation}
where $\mathbf{sentsc}(sw_{j})$ is the sentiment score of $sw_{j}$ from the sentiment lexicon and $\mathbf{worddist}(kw_i, sw_{j})$ is the number of words between $kw_i$ and $sw_{j}$ in the sentence plus one.

@@

\paragraph{Dependency distance} One way of expressing the syntactic environment of a target word with respect to a sentiment word is to determine its distance in the dependency graph. We therefore define a distance function $\mathbf{depdist}(kw_i, sw_{j})$ which returns the number of nodes in the shortest dependency path from the target word to the sentiment word in the dependency graph. The shortest path is determined using Dijkstra's shortest path algorithm \cite{Dij:59}.

\paragraph{Dependency paths}
A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the targets and sentiment words \cite{Jiang11}. The assumption would be that two words most likely are semantically related to each other if there is a meaningful grammatical relation between them. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have manually created a list of 42 grammatical dependency paths, divided into four groups, and given them a score between 1-3. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will briefly present the groups of paths and the maximum score we have assigned in each group. The paths are represented in the following format: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} indicate the direction of the traversed arc in the graph.

A first group consists of paths from subject targets to sentiment predicates. Such paths can e.g. go from a subject to a verbal predicate, subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up, or from a subject to an adjectival or nominal predicate in the context of a copular verb, subst:adj/subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn. Paths in this group can get the maximum score, 3. The combination of a subject and a predicate will result in a proposition, and we expect propositions to represent the opinion of the speaker in many cases. Also, the subject will often refer to the intentional agent of an event. If the predicate has a positive or negative sentiment, we expect that this sentiment is directed towards this intentional agent.

A second group we have considered, contains paths from subject targets to sentiment words embedded within the predicate, such as from the subject to the nominal direct object of a verb, subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn. Paths from subjects into different kinds of adverbials are also a part of this group. We consider paths from subjects to objects to be good indicators of sentiment and assign them the highest score, 3, based on the same arguments as for subject-predicate paths. Paths into adverbials get lower values, as adverbials often are less semantically connected to the predicate than objects.

The paths in our third group go from targets to sentiment words within the predicate. These include paths from nominal direct object target to verbal predicates, subst:verb{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up, and from various kinds of adverbials to verbal predicates, etc. We assume that predicate-internal paths are less good indicators of sentiment than the above groups, as such paths do not constitute a proposition. Also, arguments within the predicate usually do not represent intentional agents. Such paths will get the score 1.

Our fourth and final group of dependency paths contains paths internal to the nominal phrase, such as from target nouns to attributive adjectives, subst:adj{\textunderscore}{\textunderscore}ATR{\textunderscore}dn, and from target complements of attributive prepositions to target nouns, subst:subst{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up. A positively or negatively qualified noun will probably often represent the sentiment of the speaker. At the same time, a nominal phrase of this kind can be used in many different contexts where the holder of the sentiment is not the speaker. We assign 2 as the maximum score. Table \ref{deppath} summarizes the groups of dependency paths.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline \bf Path group & \bf Number & \bf Score range \\ \hline
Subj. to pred. & 9 & 3 \\
Subj. to pred.-internal & 13 & 3-1 \\
Pred.-internal & 6 & 1 \\
NP-internal & 14 & 2-1 \\
\hline
\end{tabular}
\end{center}
\caption{\label{deppath} Grouping of dependency paths with the number of paths and score range for each group.}
\label{tab:paths}
\end{table}

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all important dependency paths. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the dependency path, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the path. Further let $\mathbf{depdist}(kw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{depdist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependency path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{depdist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

We compare the classification performance of a set of $M$ different methods, denoted as $\Pi_1, \Pi_2, \ldots, \Pi_M$, using random effect logistic regression. Let the stochastic variable $Y_{tm} \in \{0,1\}$ represents whether method $\Pi_m, \, m \in \{1,2,\ldots,M\}$ classified the correct sentiment to sentence number $t \in \{1,2,\ldots,T\}$, where $T$ is the number of sentences in the test set. We let $Y_{tm}$ be the dependent variable of the regression model. The different methods $\Pi_1, \Pi_2, \ldots, \Pi_M$ is included as a categorical independent variable in the regression model. We also assume that classification performance of the different methods depends on the sentence to be classified, thus the sentence number is included as a random effect. The regression model is formulated as
\begin{align}
  \label{eq:5}
  \begin{split}
    P(Y_{tm} = 1) &= \text{logit}(\mu + \Pi_m + \epsilon_t + \epsilon_{tm}) \\
    P(Y_{tm} = 0) &= 1 - P(Y_{tm} = 1)
  \end{split}
 \end{align}
where $\epsilon_t$ is the sentence number random effect and $\epsilon_{tm}$ is additional random noise. Fitting the model to the observed classification performance of the different methods we are able to see if the probability of classifying correctly significantly vary between the methods.

@@

We did not find any suitable annotated text material related to political discussions in norwegian and therefore created our own. We manually selected 46 debate articles from the Norwegian online newspapers \textit{NRK Ytring, Dagbladet, Aftenposten, VG and Bergens Tidene}. To each debate article there were attached a discussion thread where readers could express their opinions and feelings towards the content of the debate article. All the text from the debate articles and the subsequent discussions were collected using text scraping \cite{Hammer13}. The debate articles were related to religion and immigration and we wanted to classify the sentiment towards all forms of the following target words: \textit{islam, muslim, quran, allah, muhammed, imam and mosque}. These are topics that typically create a lot of active discussions and disagreements.

@@

Finally the material was divided into two parts where the first half of the debate articles with subsequent discussions make up the training set and the rest constitutes a held-out test set. In the manual development of the important dependency paths, only the training set was used. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the annotation found in the corpus.
\begin{table}
  \centering
  \begin{tabular}{lcccc}
             & Negative & Neutral & Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)
  \end{tabular}
  \label{tab:1}
  \caption{Manual annotation of training and test set.}
\end{table}

@@

\section{Experiments}
\label{sec:exp}
In this study we compare four different methods based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this method WD (word distance).
\item For this method and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ which means that we do not use the important dependency paths. We denote this method A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this method CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table \ref{tab:paths}. We denote this method OD (optimal use of dependency paths)
\end{itemize}
For each method we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ tuned using the training set, and classification performance for the different methods on the test set using the parameter values tuned from the training set. The p-values are computed using the regression model presented in Section \ref{sec:sa}.
\begin{table}
  \centering
  \begin{tabular}{lcccccl}
       & $d_p$ & $d_n$ & $t_p$ & $t_n$ & Correct   & p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 145 (47\%) & \\
    A0 &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.023\\
    CW &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.024\\
    OD &    2  &   0  & 2.0  &  0.3  & 162 (53\%) & 0.016
  \end{tabular}
  \caption{The second to the fifth column show the optimal values of the parameters of the model tuned using the training set. The sixth column show the number of correct classifications and the last column shows p-values testing whether the method performs better than WD.}
  \label{tab:2}
\end{table}
We see that the sentiment shifter 'not' only has a positive effect on the classification performance when it is in front of the sentiment word. We see that using dependency distances (method A0) the classification results are significantly improved compared to using word distances in the sentence (method WD) (p-value = 0.023). Also classification based on important dependency paths (method OD) performs significantly better than WD. We also see that OD performs better than A0 (162 correct compared to 161), but this improvement is not statistically significant.

@@

Classifying sentiment in political discussions is hard because of the frequent use of irony, sarcasm and citations. In this paper we have compared the use of word distance between target word and sentiment word against metrics incorporating syntactic information. Our results show that using dependency tree distances or important dependency paths, improves the classification performance compared to using word distance.

Manually selecting important dependency paths for the aim of sentiment analysis is a hard task. A natural further step of our analysis is to expand the training and test material and use machine learning to see if there exists dependency paths that improve results compared to using dependency distance.

>>>>

\title{Sentiment classification of online political discussions: a comparison of a word-based and dependency-based method}

\author{Hugo Lewi Hammer \\
  Oslo and Akershus \\ University College \\ Department of Computer Science \\
  {\tt hugo.hammer@hioa.no} \\ \And
  Per Erik Solberg \\
   Spr\aa kbanken \\ The National Library \\ of Norway\\
  {\tt Per.Solberg@nb.no}\\ \And
  Lilja \O vrelid \\
   Department of Informatics \\ University of Oslo\\
  {\tt liljao@ifi.uio.no}}

@@

%% \end{center}
\maketitle
\begin{abstract}
Online political discussions have received a lot of attention over the past years. In this paper we compare two sentiment lexicon approaches to classify the sentiment of sentences from political discussions.
The first approach is based on applying the number of words between the target and the sentiment words to weight the sentence sentiment score. The second approach is based on using the shortest paths between target and sentiment words in a dependency graph and linguistically motivated syntactic patterns expressed as dependence paths. The methods is tested on a corpus of sentences from online Norwegian political discussions. The results show that the method based on dependency graph performs significantly better than the first approach. 
\end{abstract}

@@

%Over the past years there has been an alarming growth in hate against minorities like Muslims, Jews, Gypsies and gays in Europe, driven by right wing populism parties and extremist organizations \cite{r4,r11}. A similar increase in hate speech has been observed on the Internet \cite{r6,s2}, and experts are concerned that individuals influenced by this web content may resort to violence as a result \cite{Strommen12,Sunde13}. Hateful speech is not only observed on extremist sites, but also as comments on e.g. Twitter, YouTube and online newspaper articles.

%Social media and online discussions contain a wealth of information which can make us able to understand the extent of hate speech on the Internet. However, it turns out that academia is lacking research on social media and online radicalization \cite{s1}.

%hate and racism against is observed against minorities like Muslims, Jews, Gypsies and gays . Social media and online discussions contain a wealth of information which can make us able to understand the diversity of political opinions on the Internet. However, it turns out that academia is lacking research on social media and online radicalization \cite{s1}.

Over the past years online political discussions have received a lot of attention. E.g. the Obama 2012 election team initiated an extensive use of text analytics and machine learning techniques towards online material to guide advertising campaigns, identifying key voters, and improve fundraising \cite{Issenberg12}. There also has been a lot of concert about the alarming growth in hate and racism against minorities like Muslims, Jews and Gypsies in online discussions \cite{r6,s2}. Sentiment analysis (SA) is the discipline of automatically determining sentiment in text material and may be one important tool in understanding the diversity of opinions on the Internet. SA has mostly been used to analyze opinions in comments and reviews about commercial products, but there are also examples of SA towards political tweets and discussions, see e.g. Tumasjan et al. \shortcite{Tumasjan2010}; Chen et al. \shortcite{Chen10}. SA of political discussions is known to be a difficult task since citations, irony and sarcasm is very common \cite{Bing12}.

%Sentiment classification aims to classify a document or sentence as either positive or negative and sometimes also neutral. There are mainly two approaches, one based on machine learning and one based on based on using a list of words with given sentiment scores (lexicon-based approach). One simple lexicon-based approach is to count the number of words with positive and negative sentiment in the document as suggested by Hu and Liu \shortcite{Hu04}. One may classify the opinion of larger documents like movie or product reviews or smaller documents like tweets, comments or sentences. %% See Liu \shortcite{Bing12}, chapters three to five and references therein for the description of several opinion classification methods.

@@

%The paper is organized as follows.

@@

In this section we present two methods to classify sentences as either positive, neutral or negative towards a target word. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both target words, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $tw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the target word in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(tw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the target word appearance $tw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters are within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the sentiment towards the target word to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set, as described in section \ref{sec:exp} below.

@@

For the word distance method we use the following \textbf{imp} function
\begin{equation}
  \label{eq:2}
  \mathbf{imp}(tw_i, sw_{j}) = \frac{\mathbf{sentsc}(sw_{j})}{\mathbf{worddist}(tw_i, sw_{j})}
\end{equation}
where $\mathbf{sentsc}(sw_{j})$ is the sentiment score of $sw_{j}$ from the sentiment lexicon and $\mathbf{worddist}(tw_i, sw_{j})$ is the number of words between $tw_i$ and $sw_{j}$ in the sentence plus one.

@@

\paragraph{Dependency distance} One way of expressing the syntactic environment of a target word with respect to a sentiment word is to determine its distance in the dependency graph. We therefore define a distance function $\mathbf{depdist}(tw_i, sw_{j})$ which returns the number of nodes in the shortest dependency path from the target word to the sentiment word in the dependency graph. The shortest path is determined using Dijkstra's shortest path algorithm \cite{Dij:59}.

\paragraph{Dependency paths}
A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the targets and sentiment words \cite{Jiang11}. The assumption would be that two words most likely are semantically related to each other if there is a meaningful grammatical relation between them. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have manually created a list of 42 grammatical dependency paths, divided into four groups, and given them a score from $0-1$. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will briefly present the groups of paths and the maximum score we have assigned in each group. The paths are represented in the following format: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} indicate the direction of the traversed arc in the graph.

A first group consists of paths from subject targets to sentiment predicates. Such paths can e.g. go from a subject to a verbal predicate, subst:verb{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up, or from a subject to an adjectival or nominal predicate in the context of a copular verb, subst:adj/subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}SPRED{\textunderscore}dn. Paths in this group can get the maximum score, 1. The combination of a subject and a predicate will result in a proposition, and we expect propositions to represent the opinion of the speaker in many cases. Also, the subject will often refer to the intentional agent of an event. If the predicate has a positive or negative sentiment, we expect that this sentiment is directed towards this intentional agent.

A second group we have considered, contains paths from subject targets to sentiment words embedded within the predicate, such as from the subject to the nominal direct object of a verb, subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn. Paths from subjects into different kinds of adverbials are also a part of this group. We consider paths from subjects to objects to be good indicators of sentiment and assign them the highest score, 1, based on the same arguments as for subject-predicate paths. Paths into adverbials get lower values, as adverbials often are less semantically connected to the predicate than objects.

The paths in our third group go from targets to sentiment words within the predicate. These include paths from nominal direct object target to verbal predicates, subst:verb{\textunderscore}{\textunderscore}DOBJ{\textunderscore}up, and from various kinds of adverbials to verbal predicates, etc. We assume that predicate-internal paths are less good indicators of sentiment than the above groups, as such paths do not constitute a proposition. Also, arguments within the predicate usually do not represent intentional agents. Such paths will get the score $1/3$.

Our fourth and final group of dependency paths contains paths internal to the nominal phrase, such as from target nouns to attributive adjectives, subst:adj{\textunderscore}{\textunderscore}ATR{\textunderscore}dn, and from target complements of attributive prepositions to target nouns, subst:subst{\textunderscore}{\textunderscore}PUTFYLL{\textunderscore}up{\textunderscore}{\textunderscore}ATR{\textunderscore}up. A positively or negatively qualified noun will probably often represent the sentiment of the speaker. At the same time, a nominal phrase of this kind can be used in many different contexts where the holder of the sentiment is not the speaker. We assign $2/3$ as the maximum score. Table \ref{deppath} summarizes the groups of dependency paths.
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|cc|}
\hline \bf Path group & \bf Number & \bf Score range \\ \hline
Subj. to pred. & $9$ & $1$ \\
Subj. to pred.-internal & $13$ & $1/3-1$ \\
Pred.-internal & $6$ & $1/3$ \\
NP-internal & $14$ & $1/3-2/3$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{deppath} Grouping of dependency paths with the number of paths and score range for each group.}
\label{tab:paths}
\end{table}

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all salient dependency paths. The function $\mathbf{gram}(tw_i, sw_{j})$ returns the dependency path, and if $\mathbf{gram}(tw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(tw_i, sw_{j}) \in [0,1]$, returns the importance of the path. Further let $\mathbf{depdist}(tw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(tw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(tw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(tw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{depdist}(tw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the salient dependency path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(tw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(tw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{depdist}(tw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

@@

We did not find any suitable annotated text material related to political discussions in Norwegian and therefore created our own. We manually selected 46 debate articles from the Norwegian online newspapers \textit{NRK Ytring, Dagbladet, Aftenposten, VG and Bergens Tidene}. To each debate article there were attached a discussion thread where readers could express their opinions and feelings towards the content of the debate article. All the text from the debate articles and the subsequent discussions were collected using text scraping \cite{Hammer13}. The debate articles were related to religion and immigration and we wanted to classify the sentiment towards all forms of the following target words: \textit{islam, muslim, quran, allah, muhammed, imam and mosque}. These are topics that typically create a lot of active discussions and disagreements.

@@

Finally the material was divided into two parts where the first half of the debate articles with subsequent discussions make up the training set and the rest constitutes a held-out test set. In the manual development of the salient dependency paths, only the training set was used. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the annotation found in the corpus.
\begin{table}
  \centering
  \begin{tabular}{|l|ccc|c} 
     \hline        & \bf Negative & \bf Neutral & \bf Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)\\\hline
  \end{tabular}
  \label{tab:1}
  \caption{Manual annotation of training and test set.}
\end{table}

@@

\section{Experiments}
\label{sec:exp}

In this study we compare four different methods based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this method WD (word distance).
\item For this method and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ which means that we do not use the salient dependency paths. We denote this method A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this method CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table \ref{tab:paths}. We denote this method OD (optimal use of dependency paths)
\end{itemize}
For each method we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

We compare the classification performance of a set of $M$ different methods, denoted as $\Pi_1, \Pi_2, \ldots, \Pi_M$, using random effect logistic regression. Let the stochastic variable $Y_{tm} \in \{0,1\}$ represents whether method $\Pi_m, \, m \in \{1,2,\ldots,M\}$ classified the correct sentiment to sentence number $t \in \{1,2,\ldots,T\}$, where $T$ is the number of sentences in the test set. We let $Y_{tm}$ be the dependent variable of the regression model. The different methods $\Pi_1, \Pi_2, \ldots, \Pi_M$ is included as a categorical independent variable in the regression model. We also assume that classification performance of the different methods depends on the sentence to be classified, thus the sentence number is included as a random effect. 
%The regression model is formulated as
%\begin{align}
%  \label{eq:5}
%  \begin{split}
%    P(Y_{tm} = 1) &= \text{logit}(\mu + \Pi_m + \epsilon_t + \epsilon_{tm}) \\
%    P(Y_{tm} = 0) &= 1 - P(Y_{tm} = 1)
%  \end{split}
% \end{align}
%where $\epsilon_t$ is the sentence number random effect and $\epsilon_{tm}$ is additional random noise. 
Fitting the model to the observed classification performance of the different methods we are able to see if the probability of classifying correctly significantly vary between the methods.

\subsection{Results}

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ tuned using the training set, and classification performance for the different methods on the test set using the parameter values tuned from the training set. The p-values are computed using the regression model presented in Section \ref{sec:sa}.
\begin{table}
  \centering
  \begin{tabular}{|l|cccccl|}
\hline & $d_p$ & $d_n$ & $t_p$ & $t_n$ & \bf Accuracy   & \bf p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 47\% & \\
    A0 &    2  &   0  & 2.0  &  0.3  & 52\% & 0.023\\
    CW &    2  &   0  & 2.0  &  0.3  & 52\% & 0.024\\
    OD &    2  &   0  & 2.0  &  0.3  & 53\% & 0.016\\ \hline
  \end{tabular}
  \caption{The second to the fifth column show the optimal values of the parameters of the model tuned using the training set. The sixth column show the number of correct classifications and the last column shows p-values testing whether the method performs better than WD.}
  \label{tab:2}
\end{table}
We see that $d_n = 0$, meaning that the sentiment shifter 'not' only has a positive effect on the classification performance when it is in front of the sentiment word. We see that using dependency distances (method A0) the classification results are significantly improved compared to using word distances in the sentence (method WD) (p-value = 0.023). Also classification based on salient dependency paths (method OD) performs significantly better than WD. We also see that OD performs better than A0 (162 correct compared to 161), but this improvement is not statistically significant.

@@

Classifying sentiment in political discussions is hard because of the frequent use of irony, sarcasm and citations. In this paper we have compared the use of word distance between target word and sentiment word against metrics incorporating syntactic information. Our results show that using dependency tree distances or salient dependency paths, improves the classification performance compared to using word distance.

Manually selecting salient dependency paths for the aim of sentiment analysis is a hard task. A natural further step of our analysis is to expand the training and test material and use machine learning to see if there exists dependency paths that improve results compared to using dependency distance.
